---
title: "Graph Networks Activity"
pdf_document: default
---

### Libraries Used
```{r, message = FALSE, warning = FALSE}
library(igraph)
library(tidyverse)
library(threejs)
library(readr)
library(knitr)
library(RColorBrewer)
```

### The Data
We have provided both a custom data set and an .RDS file with a list 
outlining various cases of dialogue from the tv show 'Friends', outlining instances of speech, the speaker, and the recipient of the speech. 
```{r, message=FALSE}
node_list <- tibble(id = 0:5)
edge_list <- tibble(Root = c(0, 0, 0, 1, 1, 2, 2, 2, 3, 3, 4, 4, 5, 5, 5),
                    Destination = c(1, 2, 4, 2, 5, 0, 1, 5, 2, 4, 2, 3, 1, 2, 4))
```

```{r, message = FALSE}
friends_edgelist <- readRDS("edgelist.RDS") 
```

Then, let's view the first 10 rows of each data set:
```{r}
kable(friends_edgelist %>% head(10))

kable(edge_list %>% head(10))
```

Since the Friends network is quite a bit larger than our custom data set, 
we'll use the filter command to filter out any occurances that don't involve
the show's main characters.
```{r, message = FALSE}
friends <- c("Phoebe", "Monica", "Rachel", "Joey", "Ross", "Chandler") 
edgelist_without <- friends_edgelist %>% 
  dplyr::filter(!(from %in% friends & to %in% friends))
```

Lastly, to interact with the data most efficiently, we'll format each dataset
as a graph network; notice, each employs a different method of doing so but
arrives at the same result.

Custom data:
```{r, message = FALSE}
graph <- graph.data.frame(edge_list, directed = TRUE)
```
Friends data:
```{r, message = FALSE}
edgelist_matrix <- as.matrix(edgelist_without[ ,c("from", "to")]) 
friends_graph <- igraph::graph_from_edgelist(edgelist_matrix, directed = FALSE) %>% 
  igraph::set.edge.attribute("weight", value = edgelist_without$weight)
```

### Elements of a Graph Network

There are two main characteristics of a graph netowrk: the nodes, which are
the individual items that form the netowrk, and the edges, which signify the
connections between nodes. You can use basic commands to see the number of 
edges and vertices involved in any graph network as well as other 
characteristics such as the diameter, which is the largest number of 
connections between any two nodes.

Analysis of the custom data:
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
E(graph) # Indicates 15 edges
V(graph) # Indicates 6 nodes
get_diameter(graph) # Indicates a diameter of four

```
Analysis of the Friends data:
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
E(friends_graph) # Indicates 2961 edges signifying different instances of speech
V(friends_graph) # Indicates 650 nodes, signifying 650 different speakers
get_diameter(friends_graph) # Indicates that the greatest distance between 
#                             any characters is five degrees of separation
```
In addition to checking these basic network characteristics, certain commands 
tell you data about specific points. 

For example, the degree command can be used to tell us that Joey interacted with 348 
characters:
```{r}
degree(friends_graph, v = 'Joey')
```

You can also use this data to create visualizations of characters' social 
patterns. For example, the histogram below shows the number of times that
Chandler talked to each of his "friends", which, in this case, signifies
anyone that he spoke to more than fifteen times.
```{r message=FALSE, warning=FALSE}
chandlers_friends <- filter(friends_edgelist, from == 'Chandler') %>%
  filter(weight > 15) %>% arrange(desc(weight))

ggplot(chandlers_friends, aes(x = to, y = weight )) + geom_bar(stat = "identity", fill = "Blue") +
  ggtitle("Chandler's Friends") + xlab("Friend") + ylab("Times Spoken to") + 
  scale_fill_brewer(palette = "Blues")

```


### Visualizing the Network
Now that the basic elements of a graph network are understood, it's time to
visualize them. One way to do this is by making an ego graph, which centers 
around a single node, the ego. 
```{r}
graph_ego <- make_ego_graph(graph, 3, '0', mode = c('all'))[[1]]
fr_graph_ego <- make_ego_graph(friends_graph, 1, 'Joey', mode = c('all'))[[1]]
plot(graph_ego)
plot(fr_graph_ego)
```

As you can see, this still produces results that are not particularly helpful,
especially for the monstrous Friends netowork. At the very least, we can add 
some colors to the homemade data that indicate each node's proximity to the 
ego, number zero (the Friends network will take a little more).
```{r}
colors <- c("red", "orange","dodgerblue", "cyan")
dists <- distances(graph_ego, v = '0')
V(graph_ego)$color <- colors[dists+1]
plot(graph_ego, edge.arrow.size = .05)
```

For the Friends graph, we'll use the louvain_cluster command to organize the 
network into distinct communities, and assign those communities to the graph
object to encourage additional organization. Then, we'll create subgraphs out
of each community so they can be analyzed independently and sorted using dplyr
commands for analyzing the nodes with the most betweenness, or involvement
within the network. The resulting list will show each community's member count
and its most central member.
```{r message=FALSE, warning=FALSE, paged.print=TRUE}
# run louvain with edge weights 
louvain_partition <- igraph::cluster_louvain(friends_graph, weights = E(friends_graph)$weight) 
# assign communities to graph 
friends_graph$community <- louvain_partition$membership 
# see how many communities there are 
unique(friends_graph$community) 

communities <- data.frame() 
for (i in unique(friends_graph$community)) { 
  # create subgraphs for each community 
  subgraph <- induced_subgraph(friends_graph, v = which(friends_graph$community == i)) 
  # get size of each subgraph 
  size <- igraph::gorder(subgraph) 
  # get betweenness centrality 
  btwn <- igraph::betweenness(subgraph) 
  communities <- communities %>% 
    dplyr::bind_rows(data.frame(
      community = i, 
      n_characters = size, 
      most_important = names(which(btwn == max(btwn))) 
    ) 
    ) 
} 
knitr::kable(
  communities %>% 
    dplyr::select(community, n_characters, most_important))
```

Since three of these groups are notably smaller and led by insignificant
characters, we'll create subgraphs of each group and filter them out by
group size to ensure that only the groups lead by main characters are 
included.
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
top_five <- data.frame() 
for (i in unique(friends_graph$community)) { 
  # create subgraphs for each community 
  subgraph <- induced_subgraph(friends_graph, v = which(friends_graph$community == i)) 
  # for larger communities 
  if (igraph::gorder(subgraph) > 20) { 
    # get degree 
    degree <- igraph::degree(subgraph) 
  } 
}
```

FINALLY, it is time to visualize the network in a way that is visually pleasing
and reasonably organized. After adding some colors to represent a node's
community, scaling the nodes by degree, and adding colored edges to show the
root of the edge (or speaker of the dialogue), it's time to plot the graphs. 
Layout one, the spherical layout, is more visually pleasing but layout two
is clearly preferable as it clusters groups together and bases node placement
on community involvement and betweenness. 
```{r}
# Scaling by degree and coloring by community
V(friends_graph)$size <- 3 
V(friends_graph)$frame.color <- "white" 
V(friends_graph)$color <- friends_graph$community 
V(friends_graph)$label <- V(friends_graph)$name 
V(friends_graph)$label.cex <- 1.5 
# Coloring by speaker
edge.start <- ends(friends_graph, es = E(friends_graph), names = F)[,1] 
E(friends_graph)$color <- V(friends_graph)$color[edge.start] 
E(friends_graph)$arrow.mode <- 0 # only label central characters 
v_labels <- which(V(friends_graph)$name %in% friends) 
for (i in 1:length(V(friends_graph))) { 
  if (!(i %in% v_labels)) { V(friends_graph)$label[i] <- "" } 
}

l1 <- layout_on_sphere(friends_graph)
plot(friends_graph, rescale = T, layout = l1, main = "'Friends' Network - All Seasons")

l2 <- layout_with_mds(friends_graph) 
plot(friends_graph, rescale = T, layout = l2, main = "'Friends' Network - All Seasons")

```










